Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Squeeze

We describe our cooperative kernels programming model as an extension
to OpenCL.  However, the cooperative kernels concept is more general,
and could be applied to extend other GPU programming models, such as
CUDA and HSA.

We describe the semantics of cooperative kernels
(\mysec\ref{sec:semantics}) and summarise the important
differences between programming with regular vs.\ cooperative kernels
by showing how the motivating examples of \mysec\ref{sec:openclexamples} can be adapted to use cooperative kernels
(\mysec\ref{sec:programmingguidelines}).
We then focus on the non-functional
properties that a developer of a cooperative kernel, and the
implementer of a scheduler for cooperative kernels, should strive for
(\mysec\ref{sec:nonfunctional}).  The semantics of cooperative
kernels has been guided by the practical applications we have studied
(described in \mysec\ref{sec:portingalgorithms}), and we discuss several cases where we
might have taken different and also reasonable semantic decisions (\mysec\ref{sec:semanticalternatives}).
We conclude the section by discussing the backwards-compatible nature
of cooperative kernels with respect to blocking algorithms designed to
run on current GPU platforms
(\mysec\ref{sec:backwardscompatibility}).

@@

\subsection{Alternative Semantic Choices}\label{sec:semanticalternatives}

\ADComment{Move this to the appendix.}

@@

>>>>

We present our cooperative kernels programming model as an extension
to OpenCL.  However, the cooperative kernels concept is more general,
and could be applied to extend other GPU programming models, such as
CUDA and HSA.  We describe the semantics of the model
(\mysec\ref{sec:semantics}), use our motivating examples to discuss
programmability (\mysec\ref{sec:programmingguidelines}), outline
important nonfunctional properties that the model requires to work
successfully (\mysec\ref{sec:nonfunctional}) and explain that the
model is backwards-compatible
(\mysec\ref{sec:backwardscompatibility}).
%
In Appendices~\ref{appendix:semantics} and~\ref{appendix:semanticalternatives} (supplementary material)
we present a formal semantics for cooperative kernels atop an abstract GPU programming model,
and discuss several cases where we
might have taken different and also reasonable semantic decisions.

@@

@@

\section{Alternative Semantic Choices}\label{appendix:semanticalternatives}

The semantics of cooperative kernels has been guided by the practical
applications we have studied (described in
\mysec\ref{sec:portingalgorithms}).  We now discuss several cases
where we might have taken different and also reasonable semantic
decisions.
